Mirror, mirror on the wall. Which's the best match of them all?

### 2024-01-19 PNAS/index.Rnw
A few more edits to the conclusion

In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information. Examples include a database search or a situation similar to the example where multiple comparisons of evidence against reference samples are conducted.
%either in a database or in a series of test samples created during the examination process.
As a result, it is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions
%science grapple with this multiple comparison problem and
and that they modify lab procedures and testimony accordingly. We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.
%We have three primary suggestions for improvements targeted at making the multiple comparisons visible to everyone involved in the process of evidence evaluation.

First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence. Examiners should also report the total consecutive length or area of the recovered evidence. These pieces of information will allow for calculations of minimum and maximum number of comparisons (as we have demonstrated in this paper), and will facilitate calculation of overall error rates.

Second, researchers in this field should conduct studies relating the length/area of comparison surface to the error rate. For instance, here, we have combined studies looking at bullet striations with studies on firing pin shear marks - the striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
This was necessary, because we could not find black-box studies of wire cuts.
These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a} in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.

Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched and the number of results returned must be reported. Furthermore, the number of results used for futher manual comparison should also be reported. For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable calculation of the potential FPR.
This will help make the multiple comparison issue visible for everyone involved in evaluating the forensic evidence, from examiners themselves to jurors and judges.

>>>>

In our example we highlight a simple case where a wire  is cut using a two-sided blade, but the findings apply to many other situations where a forensic evaluation involves multiple comparisons against a set of accumulated information, such as database searches.
%either in a database or in a series of test samples created during the examination process.
It is critically important that forensic practitioners become aware of the impact of carrying out multiple comparisons on their final conclusions and that lab procedures and testimony are modified accordingly.
We suggest three strategies to improve transparency of forensic examinations and to produce more reliable estimates of examination-wide error rates.

First, examiners should report (or defense attorneys should request) the overall length or area of surfaces generated from blade cuts, test fires, and equivalent procedures intended to assess the marks made by a tool submitted for comparison with evidence. Examiners should also report the total consecutive length or area of the recovered evidence. These pieces of information will allow for calculations of minimum and maximum number of comparisons (as demonstrated in this paper), facilitating calculation of overall error rates.

Second, researchers should conduct studies relating the length/area of comparison surface to the error rate. For instance, we have pooled studies looking at bullet striations  and firing pin shear marks. The striated surfaces are of orders of magnitude different lengths, but they represent our best estimate of the error rate for striated comparisons.
This was necessary because we could not find black-box error rate studies of wire cuts.
These studies should be Type II studies as defined by Koehler \citep{koehlerIntuitiveErrorRate2017a}, in that they should be designed to assess the actual error rate when examiners are making difficult comparisons.

Finally, when databases are used at any stage of the forensic evidence evaluation process (from suitability assessment and triage to reports which will be used at trial), the number of database items searched (or comparisons made) and the number of results returned must be reported. Furthermore, the number of results used for further manual comparison should also be reported. For example, if a firearms examiner searches a local NIBIN database with 1000 entries, requests the 20 closest matches to her evidence, and then carries out a physical examination of five exemplars from the list of 20, all of those values should be clearly reported to enable calculation of the potential FPR.
This will help make the multiple comparison issue accessible to everyone involved in evaluating the value of forensic evidence, from examiners to lawyers to jurors and judges.
